---
title: "Xenium 5k Data Analysis Journey"
author: ""
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

> **Note:** 10x Genomics does not provide support for community-developed tools
> and makes no guarantees regarding their function or performance. Please contact
> tool developers with any questions.

---

# 1. Install Packages

## 1.a. Install pre-built R packages (recommended for quick setup)

```{r install-prebuilt, eval=FALSE}
system("wget https://cf.10xgenomics.com/supp/xenium/analysis-workshop/R_library_v4-5.tar.gz")
system("tar -xzvf ./R_library_v4-5.tar.gz")
system(paste0("cp -r ./R_library_v4-5/* ", .libPaths()[1]))
```

## 1.b. Install packages from R repositories (for your own computing system)

This step takes a very long time. Only use if you plan to run these analyses on
your own system. Always ensure software packages are up-to-date.

```{r install-from-repos, eval=FALSE}
lib_loc_xenium <- "./xenium_libs"
dir.create(lib_loc_xenium, showWarnings = FALSE, recursive = TRUE)

system("sudo apt-get install libmagick++-dev")
install.packages("Seurat")
install.packages("harmony")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

remotes::install_github("prabhakarlab/Banksy")
remotes::install_github("satijalab/seurat-wrappers")
remotes::install_github("immunogenomics/presto")  # speeds up DE analysis
remotes::install_github("bnprks/BPCells/r")
install.packages("hdf5r")
install.packages("sf")
install.packages("arrow")
```

---

# 2. Xenium Data Visualization and Analysis (Single Sample)

The Seurat team has developed functions to load and analyze one Xenium dataset.
`LoadXenium` handles all required files automatically.

**Required files (from your Xenium output directory):**

- `cell_feature_matrix/` — cell-level gene expression
- `cell_boundaries.parquet` — cell segmentation boundaries
- `cells.parquet` — cell-level metadata
- `transcripts.parquet` — transcript-level information

## 2.1 Download Xenium Prime Breast Cancer Data

```{r download-single-sample, eval=FALSE}
system("wget https://cf.10xgenomics.com/supp/xenium/analysis-workshop/one-sample-analysis.tar.gz")
system("tar -xzvf ./one-sample-analysis.tar.gz")
```

## 2.2 Load Xenium Data

```{r load-libraries-single}
library(Seurat)
library(sf)
library(ggplot2)
library(arrow)
library(hdf5r)
library(Banksy)
library(SeuratWrappers)
```

> **Note on `transcripts.parquet`:** From Xenium Analyzer v3.0+,
> `transcripts.csv.gz` is no longer provided. Seurat v5.2.0+ can load
> `transcripts.parquet` directly if the `arrow` package is installed
> (`install.packages("arrow")`).

**[Optional Tip] Convert `transcripts.parquet` to `transcripts.csv.gz`**
(only needed if using an older Xenium Analyzer version):

```{r convert-parquet-optional, eval=FALSE}
library(arrow)
tx <- read_parquet("./one-sample-analysis/Xenium_Prime_Breast_Cancer_FFPE_outs/transcripts.parquet")
tx$transcript_id <- as.character(tx$transcript_id)
tx_gz <- gzfile("./one-sample-analysis/Xenium_Prime_Breast_Cancer_FFPE_outs/transcripts.csv.gz", "w")
write.csv(tx, row.names = FALSE, col.names = TRUE, quote = FALSE, tx_gz)
close(tx_gz)
```

```{r load-xenium}
# Loading time is ~16 mins for this large dataset
xenium.obj <- LoadXenium(
  "./one-sample-analysis/Xenium_Prime_Breast_Cancer_FFPE_outs",
  fov          = "fov",
  segmentations = "cell",
  flip.xy      = TRUE
)

xenium.obj
```

```{r inspect-metadata}
metadata <- xenium.obj@meta.data
head(metadata, n = 3)
```

## 2.3 Transcript Plotting Functions

**Total transcripts per cell (spatial):**

```{r plot-total-transcripts, fig.width=10, fig.height=16}
ImageFeaturePlot(
  xenium.obj, fov      = "fov",
  features  = "nCount_Xenium",
  max.cutoff = "q95"
) + scale_y_reverse()
```

**Spatial gene expression — selected genes:**

```{r plot-gene-expression, fig.width=5, fig.height=7}
ImageFeaturePlot(xenium.obj, features = "ERBB2",
                 max.cutoff = "q95", cols = c("black", "white")) +
  scale_y_reverse()

ImageFeaturePlot(xenium.obj, features = "IL7R",
                 max.cutoff = "q95", cols = c("black", "white")) +
  scale_y_reverse()
```

## 2.4 Select Sub-Region of Interest

### 2.4.1 Select Sub-Region by Coordinates

Export coordinates from Xenium Explorer. The CSV will look like:

```
# Selection name: Selection 1
# Area (µm^2): 6238188.5
X,Y
2082.27,2883.95
4605.83,2883.95
4605.83,5355.93
2082.27,5355.93
2082.27,2883.95
```

```{r crop-by-coordinates}
min_x <- 2082.2669778917857
min_y <- 2883.948207577963
max_x <- 4605.827306527083
max_y <- 5355.927347229994

options(future.globals.maxSize = 1e10)
crop <- Crop(xenium.obj[["fov"]],
             x = c(min_x, max_x),
             y = c(min_y, max_y))

xenium.obj.new <- xenium.obj
xenium.obj.new[["myarea"]] <- crop
xenium.obj.new
```

```{r plot-cropped-region, fig.width=8, fig.height=8}
ImageFeaturePlot(
  xenium.obj.new, fov       = "myarea",
  features  = "nCount_Xenium",
  cols      = c("black", "white"),
  max.cutoff = "q95"
) + scale_y_reverse()
```

### 2.4.2 Select Sub-Region by Cell IDs

Export a cell stats CSV from Xenium Explorer, then:

```{r select-by-cell-ids}
select_cells <- read.csv(
  "./one-sample-analysis/Selection_1_cells_stats.csv",
  skip = 2, header = TRUE
)
head(select_cells, n = 3)

xenium.obj.sub <- subset(xenium.obj, cells = select_cells$Cell.ID)
xenium.obj.sub
```

```{r plot-subset, fig.width=8, fig.height=8}
ImageFeaturePlot(
  xenium.obj.sub, fov       = "fov",
  features  = "nCount_Xenium",
  cols      = c("black", "white"),
  max.cutoff = "q95"
) + scale_y_reverse()
```

## 2.5 Downstream Analysis

### 2.5.1 Quality Control and Preprocessing

```{r qc-histogram}
hist(xenium.obj$nCount_Xenium, xlim = c(0, 200), breaks = 5000,
     main = "Transcript counts per cell", xlab = "nCount_Xenium")
abline(v = 20, col = "red")
```

```{r qc-filter}
# Upper limit: 98th percentile; Lower limit: 20 (set from histogram above)
thres <- quantile(xenium.obj$nCount_Xenium, 0.98)
xenium.obj <- subset(xenium.obj,
                     subset = nCount_Xenium >= 20 & nCount_Xenium <= thres)
```

```{r normalize}
# Library-size normalization; scale factor = median transcript count
xenium.obj <- NormalizeData(xenium.obj,
                            scale.factor = median(xenium.obj$nCount_Xenium))
```

```{r variable-features}
# For Xenium Prime (>5k genes): top 2k variable genes
xenium.obj <- FindVariableFeatures(xenium.obj)

# For Xenium v1 panel (use ALL genes instead):
# xenium.obj <- FindVariableFeatures(xenium.obj, nfeatures = nrow(xenium.obj))
```

```{r scale-pca}
xenium.obj <- ScaleData(xenium.obj)
xenium.obj <- RunPCA(xenium.obj, verbose = FALSE)
```

```{r cluster-umap}
xenium.obj <- FindNeighbors(xenium.obj, dims = 1:30)
xenium.obj <- FindClusters(xenium.obj, verbose = FALSE)
xenium.obj <- RunUMAP(xenium.obj, dims = 1:30, verbose = FALSE)
```

### 2.5.2 UMAP of Clustering Results

```{r umap-plot}
DimPlot(xenium.obj)
```

### 2.5.3 Spatial Clustering Plot

```{r spatial-cluster-plot, fig.width=10, fig.height=14}
ImageDimPlot(xenium.obj) + scale_y_reverse()
```

### 2.5.4 Export Clustering Results for Xenium Explorer

```{r export-clustering}
clustering_res <- FetchData(xenium.obj, vars = "seurat_clusters")
clustering_res$cell_id <- rownames(clustering_res)
clustering_res <- clustering_res[, c("cell_id", "seurat_clusters")]
colnames(clustering_res) <- c("cell_id", "group")

head(clustering_res, n = 3)

write.table(clustering_res, "seurat_clustering_results.csv",
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = ",")
```

### 2.5.5 Find Marker Genes per Cluster

```{r find-markers}
all.markers <- FindAllMarkers(xenium.obj, only.pos = TRUE,
                              min.pct = 0.05, logfc.threshold = 1)
head(all.markers)

write.table(all.markers, "marker_genes_in_seurat_clusters.csv",
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = ",")
```

### 2.5.6 Cell Annotation Based on Marker Genes

Suggested workflow:

1. Get marker genes per cluster from `marker_genes_in_seurat_clusters.csv`
2. Use a tool like **[Enrichr](https://maayanlab.cloud/Enrichr/)** to find
   associated cell types
3. Cross-reference spatial location in Xenium Explorer (with H&E if available)

**Example — verify macrophage marker C1QC in cluster 6:**

```{r vln-macrophage, fig.width=12, fig.height=8}
VlnPlot(xenium.obj, features = "C1QC", layer = "data", pt.size = 0)
```

---

# 3. Multi-Sample Integration (Seurat Sketch + Harmony)

Two Xenium Prime datasets are integrated here:

- **Human breast cancer** Xenium Prime
- **Human cervical cancer** Xenium Prime

For large datasets (>1.5M cells combined), we use **Seurat sketch** to
subsample cells while preserving rare populations, and **Harmony** for batch
correction.

Only `cell_feature_matrix.h5` is required per sample.

## 3.1 Load Libraries

```{r load-libraries-multi}
library(Seurat)
library(ggplot2)
library(BPCells)
library(SeuratWrappers)
options(future.globals.maxSize = 1e10)
```

## 3.2 Download Data

```{r download-multi-sample, eval=FALSE}
system("wget https://cf.10xgenomics.com/supp/xenium/analysis-workshop/multi-sample-analysis.tar.gz")
system("tar -xzvf ./multi-sample-analysis.tar.gz")
```

## 3.3 Create On-Disk Matrix Representations

```{r create-ondisk-breast}
xenium.breast <- open_matrix_10x_hdf5(
  path         = "./multi-sample-analysis/human_breast_xenium/cell_feature_matrix.h5",
  feature_type = "Gene Expression"
)
write_matrix_dir(mat = xenium.breast, dir = "./xenium_breast_matrix", overwrite = TRUE)
```

```{r create-ondisk-cervical}
xenium.cervical <- open_matrix_10x_hdf5(
  path         = "./multi-sample-analysis/human_cervical_xenium/cell_feature_matrix.h5",
  feature_type = "Gene Expression"
)
write_matrix_dir(mat = xenium.cervical, dir = "./xenium_cervical_matrix", overwrite = TRUE)
```

```{r load-ondisk-and-rename}
# Load on-disk matrices
xenium.breast.count   <- open_matrix_dir(dir = "./xenium_breast_matrix")
xenium.cervical.count <- open_matrix_dir(dir = "./xenium_cervical_matrix")

# Make cell IDs unique across samples
colnames(xenium.breast.count)   <- paste(colnames(xenium.breast.count),   "breast",   sep = "_")
colnames(xenium.cervical.count) <- paste(colnames(xenium.cervical.count), "cervical", sep = "_")
```

```{r convert-ensembl-breast}
# Convert Ensembl IDs to gene names (breast)
features <- read.table(
  "./multi-sample-analysis/human_breast_xenium/features.tsv.gz",
  sep = "\t", header = FALSE
)
features <- features[features$V3 == "Gene Expression", ]
features <- features[match(rownames(xenium.breast.count), features$V1), ]
rownames(xenium.breast.count) <- features$V2
```

```{r convert-ensembl-cervical}
# Convert Ensembl IDs to gene names (cervical)
features <- read.table(
  "./multi-sample-analysis/human_cervical_xenium/features.tsv.gz",
  sep = "\t", header = FALSE
)
features <- features[features$V3 == "Gene Expression", ]
features <- features[match(rownames(xenium.cervical.count), features$V1), ]
rownames(xenium.cervical.count) <- features$V2
```

**[Tip] Working with multiple tissues from one Xenium slide (e.g., TMA)?**

Use Section 2.4.2 to get cell IDs per tissue region, then subset the matrix:

```{r tma-subset-tip, eval=FALSE}
gene_exp_data <- Read10X_h5("./multi-sample-analysis/human_breast_xenium/cell_feature_matrix.h5")
gene_exp_data <- gene_exp_data$`Gene Expression`

tissue1_cells <- read.csv("./multi-sample-analysis/Tissue_1_cells_stats.csv",
                          skip = 2, header = TRUE)

gene_exp_data_tissue1 <- gene_exp_data[, tissue1_cells$Cell.ID]
write_matrix_dir(mat = gene_exp_data_tissue1, dir = "./tissue1_matrix", overwrite = TRUE)

tissue1.mat <- open_matrix_dir(dir = "./tissue1_matrix")
colnames(tissue1.mat) <- paste(colnames(tissue1.mat), "tissue1", sep = "_")
# Repeat for tissues 2, 3, etc., then proceed to Section 3.4
```

## 3.4 Create Merged Seurat Object

```{r create-seurat-merged}
countlist <- list(xenium.breast.count, xenium.cervical.count)
names(countlist) <- c("breast", "cervical")

merge.obj <- CreateSeuratObject(countlist)

# Tag each cell with its sample origin
samples <- sapply(strsplit(colnames(merge.obj), "_"), function(x) x[2])
merge.obj$samples <- samples
```

## 3.5 QC and Sketch Subsampling

```{r qc-multi-histogram}
hist(merge.obj$nCount_RNA, xlim = c(0, 200), breaks = 5000,
     main = "Transcript counts per cell", xlab = "nCount_RNA")
abline(v = 20, col = "red")
```

```{r qc-multi-filter}
thres <- quantile(merge.obj$nCount_RNA, 0.98)
merge.obj <- subset(merge.obj,
                    subset = nCount_RNA >= 20 & nCount_RNA <= thres)
```

```{r normalize-multi}
merge.obj <- NormalizeData(merge.obj,
                           scale.factor = median(merge.obj$nCount_RNA))
```

```{r variable-features-multi}
# For Xenium Prime (>5k genes):
merge.obj <- FindVariableFeatures(merge.obj, verbose = FALSE)

# For Xenium v1 panel (use ALL genes):
# merge.obj <- FindVariableFeatures(merge.obj, nfeatures = nrow(merge.obj))
```

```{r sketch-subsample}
# 25k cells sampled per dataset (50k total in memory)
# Adjust ncells based on your dataset sizes
# e.g., ncells = c(25000, 35000) for unequal sampling
merge.obj <- SketchData(
  merge.obj,
  ncells         = 25000,
  method         = "LeverageScore",
  sketched.assay = "sketch",
  features       = VariableFeatures(merge.obj)
)

merge.obj
```

## 3.6 Conventional Processing on Sketch (50k cells)

```{r process-sketch}
DefaultAssay(merge.obj) <- "sketch"
merge.obj <- FindVariableFeatures(merge.obj, verbose = FALSE)
merge.obj <- ScaleData(merge.obj, verbose = FALSE)
merge.obj <- RunPCA(merge.obj, verbose = FALSE)
```

## 3.7 Batch Correction with Harmony

**First, check for batch effects without correction:**

```{r umap-no-correction}
merge.obj <- RunUMAP(merge.obj, dims = 1:30, return.model = TRUE, verbose = FALSE)
DimPlot(merge.obj, group.by = "samples")
```

**Apply Harmony batch correction:**

```{r harmony-integration}
merge.obj <- IntegrateLayers(
  merge.obj,
  method        = HarmonyIntegration,
  orig          = "pca",
  new.reduction = "harmony",
  dims          = 1:30
)
```

```{r umap-harmony}
merge.obj <- RunUMAP(merge.obj, reduction = "harmony", dims = 1:30,
                     return.model = TRUE, verbose = FALSE)
```

```{r cluster-harmony}
merge.obj <- FindNeighbors(merge.obj, reduction = "harmony", dims = 1:30)
merge.obj <- FindClusters(merge.obj, resolution = 1, verbose = FALSE)
```

```{r plot-harmony-umap}
DimPlot(merge.obj, group.by = "samples")
DimPlot(merge.obj)
```

## 3.8 Project Results from 50k Sketch to All Cells

```{r project-integration}
merge.obj <- ProjectIntegration(
  object         = merge.obj,
  sketched.assay = "sketch",
  assay          = "RNA",
  reduction      = "harmony"
)
```

```{r project-data}
merge.obj <- ProjectData(
  object             = merge.obj,
  assay              = "RNA",
  full.reduction     = "harmony.full",
  sketched.assay     = "sketch",
  sketched.reduction = "harmony.full",
  umap.model         = "umap",
  dims               = 1:30,
  refdata            = list(cluster_full = "seurat_clusters")
)
```

```{r plot-full-umap}
DimPlot(merge.obj, reduction = "full.umap", group.by = "cluster_full", label = TRUE)
DimPlot(merge.obj, reduction = "full.umap", group.by = "samples")
```

## 3.9 Export Clustering Results for Xenium Explorer

```{r export-multi-clustering}
clusters_all <- FetchData(merge.obj, vars = c("samples", "cluster_full"),
                          layer = "counts")
clusters_all$cell_id <- rownames(clusters_all)

# Breast cancer clustering
clusters_all_breast <- clusters_all[
  clusters_all$samples == "breast", c("cell_id", "cluster_full")
]
colnames(clusters_all_breast) <- c("cell_id", "group")
write.table(clusters_all_breast, "integrated_clusters_for_breast.csv",
            sep = ",", col.names = TRUE, row.names = FALSE, quote = FALSE)

# Cervical cancer clustering
clusters_all_cervical <- clusters_all[
  clusters_all$samples == "cervical", c("cell_id", "cluster_full")
]
colnames(clusters_all_cervical) <- c("cell_id", "group")
write.table(clusters_all_cervical, "integrated_clusters_for_cervical.csv",
            sep = ",", col.names = TRUE, row.names = FALSE, quote = FALSE)
```

## 3.10 Find Cluster Marker Genes

```{r find-markers-multi}
DefaultAssay(merge.obj) <- "sketch"
merge.obj[["sketch"]] <- JoinLayers(merge.obj[["sketch"]])

all.markers <- FindAllMarkers(merge.obj, only.pos = TRUE)

write.table(all.markers, "markers_for_all_clusters_in_sketch.csv",
            sep = ",", col.names = TRUE, row.names = FALSE, quote = FALSE)
```

**Example: Visualize macrophage marker C1QC across clusters (cluster 6 = macrophage):**

```{r vln-macrophage-multi, fig.width=12, fig.height=8}
VlnPlot(merge.obj, assay = "RNA", features = "C1QC",
        layer = "data", pt.size = 0, group.by = "cluster_full")
```

**[Tip] Save and reload the integrated Seurat object:**

```{r save-load-seurat, eval=FALSE}
# Save — keep this .rds file together with xenium_breast_matrix/ and xenium_cervical_matrix/
saveRDS(merge.obj, "Xenium_integrated_Seurat.rds")

# Reload (on-disk folders must be present)
merge.obj <- readRDS("./Xenium_integrated_Seurat.rds")
```

## 3.11 Subset and Analyze Macrophages

```{r subset-macrophages}
macrophage_cells <- names(merge.obj$cluster_full[merge.obj$cluster_full == 6])
macro.obj <- subset(merge.obj, cells = macrophage_cells)
macro.obj
```

```{r process-macrophages}
DefaultAssay(macro.obj) <- "RNA"

# For Xenium Prime (>5k genes):
macro.obj <- FindVariableFeatures(macro.obj, verbose = FALSE)
# For Xenium v1 panel: macro.obj <- FindVariableFeatures(macro.obj, nfeatures = nrow(macro.obj))

macro.obj <- ScaleData(macro.obj, verbose = FALSE)
macro.obj <- RunPCA(macro.obj, verbose = FALSE)
```

```{r harmony-macrophages}
macro.obj <- IntegrateLayers(
  macro.obj,
  method        = HarmonyIntegration,
  orig          = "pca",
  new.reduction = "harmony.full",
  dims          = 1:30
)
```

```{r umap-macrophages, fig.width=7, fig.height=7}
macro.obj <- RunUMAP(macro.obj, reduction = "harmony.full", dims = 1:30,
                     reduction.name = "umap.full",
                     reduction.key  = "UMAP_full_", verbose = FALSE)

DimPlot(macro.obj, reduction = "umap.full", group.by = "samples")
```

**Differential gene expression: breast vs. cervical macrophages:**

> ⚠️ **Statistical note:** p-values below should be interpreted with caution —
> each cell is treated as an independent replicate. For robust DEG analysis,
> use biological replicates (≥3 per condition) with a pseudo-bulk approach.
> See: [Seurat pseudo-bulk DE tutorial](https://satijalab.org/seurat/articles/de_vignette).

```{r deg-macrophages}
Idents(macro.obj) <- "samples"
macro.obj[["RNA"]] <- JoinLayers(macro.obj[["RNA"]])

macr.marker <- FindMarkers(macro.obj, ident.1 = "breast", ident.2 = "cervical")

macr.marker$gene <- rownames(macr.marker)
write.table(macr.marker, "macrophage_DGE.csv",
            sep = ",", col.names = TRUE, row.names = FALSE, quote = FALSE)

head(macr.marker)
```

```{r plot-ctsl-feature, fig.width=8, fig.height=8}
FeaturePlot(macro.obj, features = "CTSL", reduction = "umap.full")
```

```{r plot-ctsl-violin, fig.width=8, fig.height=8}
VlnPlot(macro.obj, "CTSL", group.by = "samples")
```

Higher CTSL expression in macrophages is associated with tumor progression and
poor prognosis (Zhang et al., 2024; Alderton, 2010; Xiao et al., 2024). The
analysis shows elevated CTSL in breast cancer macrophages vs. cervical cancer,
suggesting a more pro-tumor macrophage role in breast cancer.

---

# 4. Spatially-Informed Clustering with Banksy

Standard clustering uses only each cell's gene expression. **Banksy** also
incorporates mean expression and expression gradients from neighboring cells,
improving cell-type resolution in spatial data.

Banksy was ranked among the top-performing spatial clustering tools with
strong scalability (Hu et al., 2024; Singhal et al., 2024).

**Required files (same as Section 2):**

- `cell_feature_matrix/`
- `cell_boundaries.parquet`
- `cells.parquet`
- `transcripts.parquet`

## 4.1 Data Download and Preprocessing

```{r banksy-setup}
system("sudo apt-get install libmagick++-dev")

library(Seurat)
library(sf)
library(ggplot2)
library(Banksy)
library(SeuratWrappers)
```

```{r download-banksy-data, eval=FALSE}
system("wget https://cf.10xgenomics.com/supp/xenium/analysis-workshop/one-sample-analysis.tar.gz")
system("tar -xzvf ./one-sample-analysis.tar.gz")
```

```{r load-xenium-banksy}
xenium.obj <- LoadXenium(
  "./one-sample-analysis/Xenium_Prime_Breast_Cancer_FFPE_outs",
  fov           = "fov",
  segmentations = "cell",
  flip.xy       = TRUE
)
```

```{r qc-banksy}
hist(xenium.obj$nCount_Xenium, xlim = c(0, 200), breaks = 5000,
     main = "Transcript counts per cell", xlab = "nCount_Xenium")
abline(v = 20, col = "red")

thres <- quantile(xenium.obj$nCount_Xenium, 0.98)
xenium.obj <- subset(xenium.obj,
                     subset = nCount_Xenium >= 20 & nCount_Xenium <= thres)
```

```{r preprocess-banksy}
xenium.obj <- NormalizeData(xenium.obj,
                            scale.factor = median(xenium.obj$nCount_Xenium))

# For Xenium Prime (>5k genes):
xenium.obj <- FindVariableFeatures(xenium.obj)
# For Xenium v1: xenium.obj <- FindVariableFeatures(xenium.obj, nfeatures = nrow(xenium.obj))

xenium.obj <- ScaleData(xenium.obj)
```

## 4.2 Run Banksy

- **`lambda = 0.2`** — recommended for cell typing (controls neighborhood weight;
  `lambda = 0` uses only each cell's own expression)
- **`k_geom = 15`** — number of neighboring cells to define the neighborhood

```{r run-banksy}
# ~11 mins for a 490k-cell Xenium Prime dataset
xenium.obj <- RunBanksy(
  xenium.obj,
  lambda   = 0.2,
  k_geom   = 15,
  assay    = "Xenium",
  slot     = "data",
  features = "variable",
  verbose  = FALSE
)
```

```{r banksy-pca-cluster-umap}
xenium.obj <- RunPCA(xenium.obj, assay = "BANKSY",
                     features = rownames(xenium.obj), npcs = 30)

xenium.obj <- FindNeighbors(xenium.obj, dims = 1:30)
xenium.obj <- FindClusters(xenium.obj, resolution = 0.5)
xenium.obj <- RunUMAP(xenium.obj, dims = 1:30, verbose = FALSE)

DimPlot(xenium.obj)
```

## 4.3 Export Banksy Clustering Results for Xenium Explorer

```{r export-banksy}
clustering_res <- FetchData(xenium.obj, vars = "seurat_clusters")
clustering_res$cell_id <- rownames(clustering_res)
clustering_res <- clustering_res[, c("cell_id", "seurat_clusters")]
colnames(clustering_res) <- c("cell_id", "group")

head(clustering_res, n = 3)

write.table(clustering_res, "banksy_clustering_results.csv",
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = ",")
```

---

*End of Xenium 5k Data Analysis Notebook*
